---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyrgee

<!-- badges: start -->
[![R-CMD-check](https://github.com/r-tidy-remote-sensing/tidyrgee/workflows/R-CMD-check/badge.svg)](https://github.com/r-tidy-remote-sensing/tidyrgee/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/tidyrgee)](https://CRAN.R-project.org/package=tidyrgee)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of tidyrgee is to create tidyverse style methods (filter, group_by, summarise, etc) for Google Earth Engine (GEE) Images and ImageCollections. 

## Installation

You can install the development version of tidyrgee from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("r-tidy-remote-sensing/tidyrgee")
```

## Example

This is a basic example which shows you how to solve a common problem: filter an ImageCollection by date and then group by year and month and then take the mean.  

### load and initialize
```{r example}

library(tidyrgee)
library(rgee)
ee_Initialize()
```

### dplyresque syntax

In this example, by piping in  `filter`, `group_by` , `summarise`  just as you would for a data.frame you are sequentially:

1. filtering the ImageCollection by date
2. grouping it by year
3. summarizing at the pixel-level by taking the mean of all pixels in your date range for each year

The result will be an `ImageCollection` with the one `Image` per year that holds the mean pixel values

```{r}
modis_ic <- ee$ImageCollection("MODIS/006/MOD13Q1")

modis_mean_ndvi_yearly <- modis_ic$select("NDVI") %>% # still rgee/GEE syntax
   filter(date>="2016-01-01",date<="2019-12-31") %>%
   group_by(year) %>%
   summarise(stat="mean")
```

