---
title: "ab-tidyee-class-framework"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ab-tidyee-class-framework}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyrgee)
library(rgee)
library(dplyr)
ee_Initialize()

```

## Intro

In addition to building nice dplyresque/tidy syntax wrappers around `rgee/GEE` functions. We have decided to explore the possibility of introducing a new framework which includes a new class object: "tidyee". 

To use this framework your `ImageCollection` or `Image` has to be converted to `tidyee` class using the new `as_tidy_ee` function as shown below:

```{r}
modis_ic <- ee$ImageCollection("MODIS/006/MOD13Q1")
modis_ic_tidy <- as_tidyee(modis_ic)
```

As you can see below the new object (`modis_ic_tidy`) is a named list (of class "tidyee") containing :

1. `ee_ob`: the original ee_object (in this case `ImageCollection`)
2. `vrt`: virtual table holding key properties of the original ee_object

```{r}
modis_ic_tidy$ee_ob
modis_ic_tidy$vrt
```

The virtual table (`vrt`) data.frame allows us to leverage all the power and functionality of dplyr to filter, mutate,group, etc. An S3 class method `filter.tidyee` has been written which, essentially, first filters the `vrt` based on conditions supplied to the `filter` argument and then uses the filtered data.frame (vrt) to filter/subset the `ImageCollection`. The vrt comes with several pre-defined  columns useful for filtering(date, year, month), but `mutate` can be used to add any new columns/categories for filtering.  

Below is an example using months to filter

```{r}
# library(tidyverse)
# library(tidyrgee)
# filter |> debugonce()

modis_march_april <- modis_ic_tidy |> 
  filter(month %in% c(3,4))




```

Now we show an example of mutating a new category and then filtering the `tidyee` by that column
```{r}

modis_filt_growing_season <- modis_ic_tidy |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  filter(crop_cycle=="planting")


modis_split_crop_cycle <- modis_ic_tidy |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  group_by(crop_cycle) |> 
  group_split()

modis_split_crop_cycle[[2]]$ee_ob$aggregate_array("tidyee_index")$getInfo()
modis_split_crop_cycle[[2]]$ee_ob$aggregate_array("system:time_start")$getInfo()
modis_split_crop_cycle[[2]]$ee_ob$size()$getInfo()


external_group_output <- modis_ic_tidy |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  group_by(crop_cycle) |> 
  summarise(
    stat="mean",join_bands=F
  )


external_group_output$ee_ob$size()$getInfo()

external_group_output_multi <- modis_ic_tidy |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  group_by(crop_cycle) |> 
  summarise(
    stat=list("mean","sd","min"),join_bands=F
  )


external_group_output_multi_joined <- modis_ic_tidy |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  group_by(crop_cycle) |> 
  summarise(
    stat=list("mean","sd")
  )

```

Here we show how you can perform pixel-level summary statistics with `summarise` function. This is typically referred to as `compositing` in GEE documentation as well as other remote sensing literature.

```{r}

modis_mean_by_yrmo <- modis_ic_tidy |> 
  group_by(year,month) |> 
  summarise(stat = "mean")

modis_mean_by_yrmo <- modis_ic_tidy |> 
  select(NDVI,EVI) |> 
  group_by(year,month) |> 
  summarise(stat = "mean")
```

It's nice that you can summarise by multiple different statistics at once if you want.
```{r}
modis_mean_and_sd_by_yrmo <- modis_ic_tidy |> 
  select("NDVI") |> 
  group_by(year,month) |> 
  summarise(stat = list("mean","sd"))

modis_mean_and_sd_by_yrmo
```

Next we will show how you can mutate a new category and then group by and summarise to that category. This is nice and seems to be working well. However, we still need to work out to deal with disappearing attributes after running dplyr verbs like `group_split`. This does not seem to affect the results, but just the print methods. Could be a potential solution in `vctrs` package. However, it might make more sense to just store `band_names` as col/list-col instead of attribute.

```{r}
modis_ic_tidy |> 
  # select("NDVI") |> 
  mutate(crop_cycle= case_when( 
    month %in% c(4,5)~ "land prep",
    month %in% c(6,7)~ "planting",
    month %in% c(8,9,10)~"growing",
    month ==11~ "harvesting",
    TRUE ~ "other"
      )
           ) |> 
  group_by(crop_cycle) |> 
  summarise(
    stat="mean"
  )

```

select & inner_join example
```{r}

modis_monthly_baseline_mean <- modis_ic_tidy |> 
  select("NDVI") |> 
  filter(year %in% 2000:2015) |> 
  group_by(month) |> 
  summarise(stat="mean")

modis_monthly_baseline_sd <- modis_ic_tidy |> 
  select("NDVI") |> 
  filter(year %in% 2000:2015) |> 
  group_by(month) |> 
  summarise(stat="sd")

modis_monthly_baseline <- modis_monthly_baseline_mean |> 
  inner_join(modis_monthly_baseline_sd, by="month")

modis_monthly_baseline

```

```{r,eval =F}

point_sample_buffered <- tidyrgee::bgd_msna |> 
  dplyr::sample_n(3) |> 
  sf::st_as_sf(coords=c("_gps_reading_longitude",
                        "_gps_reading_latitude"), crs=4326) |>
  sf::st_transform(crs=32646) |> 
  sf::st_buffer(dist = 500) |> 
  dplyr::select(`_uuid`) 


ndvi_monthly_mean_at_pt<- modis_monthly_baseline_mean |> 
  ee_extract_tidy(y = point_sample_buffered, 
             fun="mean",
             scale = 500)

# just to show that it also works on imageCollection
modis_monthly_baseline_ic<- modis_monthly_baseline_mean |> 
  as_ee()

modis_monthly_baseline_ic |> 
  ee_extract_tidy(y = point_sample_buffered, 
             fun="mean",
             scale = 500)

# and image
modis_monthly_baseline_img_first <-  modis_monthly_baseline_ic$first()

modis_monthly_baseline_img_first |> 
  ee_extract_tidy(y = point_sample_buffered, 
             fun="mean",
             scale = 500)


```


## Limitations/Next Steps

Below I list properties of this approach that could be considered potential downsides and list potential ways to circumvent or minimize these downsides.

**1. The new `tidyee` object reduces interoperability with `rgee`**

Some potential ideas to improve:

a. maybe very simple functions to switch resulting `tidyee` object back to `ee$ImageCollection` or `ee$Image` (maybe `as_ic`, `as_img`,`as_ee`)
b.add option to make `tidyee` on fly from `ee$Image`/`ee$ImageCollection` and then also include something like `return_ic`as a logical switch which will just return `ee$Image` or `ee$ImageCollection` instead of `tidyee` class.

So far I lean towards **a** and just create a `as_ee` function to implement it.


```{r}
modis_ic_tidy |> 
  as_ee()

```

**2. as_tidyee makes the process take slightly longer**

a. Since `as_tidyee` relies on client-side operation (primarily `rgee::ee_get_date_ic`) this function requires some start-up time investment. However, I am thinking that this one-time investment will actually save time since when using the `tidyee` object we will have constantly updating data.frame which is basically updated instantaneously as we filter and process the `ImageCollection`. This could allow nice print methods and querying without having to perform the `rgee`/client side functions like `rgee::ee_print` and `getInfo` repeatedly in work-flows which take just as much time as `as_tidyee` every time they are run.
b. To make sure these percieved benefits are actual benefits we should: 1) include checks/assertions at the end of each process to ensure the `ee_ob` and `vrt` are in perfect agreement, 2) think about including more information (bands, properties) in the the print methods for tidyee


Might be worth prefixing `dplyr` functions with `ee_` to avoid conflicts?

```{r}

modis_ndvi_baseline <- modis_ic_tidy |>
  select("NDVI") |>
  filter(year %in% c(2000:2015)) |> 
  group_by(month) |>
  summarise(stat = list("mean","sd"))

modis_ndvi_current <- modis_ic_tidy |>
  select("NDVI") |>
  filter(year %in% c(2022)) |> 
  group_by(month) |>
  summarise(stat = "mean")

modis_mean_and_sd_by_yrmo


modis_monthly_baseline_current <- modis_ndvi_baseline |> 
  inner_join(modis_ndvi_current |> select(NDVI_mean_current="NDVI"), by="month")

modis_current_with_baseline <- modis_ndvi_current |>
  select(NDVI_mean_current="NDVI") |> 
  inner_join(modis_ndvi_baseline, by="month")

```



