---
title: "ac-summarise-purrring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ac-summarise-purrring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Setup
```{r setup}

library(rgee)
library(tidyrgee)
ee_Initialize()

modis_link <- "MODIS/006/MOD13Q1"
modisIC <- ee$ImageCollection(modis_link)
modis_ndvi_tidy <- as_tidyee(modisIC) |> 
  select("NDVI")
```

# Run summarise

Currently it's now working with multiples statistics. I did a few changes:

- I made an `inner_join` method for both `tidyee` and `ImageCollection`.  The one being utilized here inside of summarise is the `tidyee` method
- I made the last step of the `inner_join` function `create_tidyee(..., tidyee_index=F)`. The **tidyee_index=F** seems to be the crucial to make this run without getting hung up giving the weird python "time has elapsed" error

```{r}
summarised_mean_sd <- modis_ndvi_tidy |> 
  filter(year %in% 2000:2015) |> 
  group_by(month) |> 
  summarise(stat=list("mean","sd","min"))
```

- it is very odd why creating the index (with setting `tidyee=T`) causes this issue because **as you can see in the following chunk** I can easily use the `set_idx` function and have it run with no delay

```{r}
ic_with_tidyee_idx <- summarised_mean_sd$ee_ob |> 
  tidyrgee:::set_idx()
ic_with_tidyee_idx$aggregate_array("tidyee_index")$getInfo()
```

- Furthermore, after creating the summarised merged/joined tidyee imageCollection/vrt, I can run `create_tidyee` with `tidyee_index=T` and have it run instantly. Unforutunately no matter how i put this at the end of the `summarise` function it hangs.... WHY!?  

```{r}
create_tidyee(x = summarised_mean_sd$ee_ob ,vrt = summarised_mean_sd$vrt,tidyee_index = T)

```
